### 2011-04-18 Text/paper.tex
Finished the preliminary version of the paper.

\documentclass{article}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{bm}
\usepackage{color}
\usepackage{tabularx}
\begin{document}
\begin{abstract}
Classical Bretherton problem describes the propagation of gas fingers through liquid media with
thin liquid films between bubbles and the channel walls. The bubble shape and flow patterns are
complicated functions of the capilary number $Ca$ and Reynolds number $Re$. Recently we investigated
the applicability and parameters choice for the two-dimensional case Bretherton problem (flow
between plates) using the free-energy binary liquid lattice Boltzmann method (LBM)
\cite{kuzmin-binary2d}. This
work is the continuation of the previous work. It is focused on the three-dimensional capillaries
simulations with square crosssections to
validate the binary liquid LBM for  Bretherton phenomena in the moderate range of the capillary
number $0.1\leq Ca \leq 1.0$.  
The flow is driven by a body force,
and
periodic boundary conditions are applied in the streamwise direction. The results show that the
binary liquid model is able to capture a number of phenomena happening in three-dimensional
capillaries, as the existence of the vortex in front of the bubble and the bubble radii dependency
on the capillary number. Therefore, the lattice Boltzmann free energy binary liquid model can be
used to simulate the Bretherton problem with good accuracy. 
\end{abstract}

\section{Introduction}
The Taylor/Bretherton \cite{bretherton} flow deals with long gas bubbles moving through liquid in
narrow channels. Depending on the geometry it was found that the deposited film thickness
is a complicated function of the capillary number $Ca$:
\begin{equation}
\label{capillary:number:definition}
Ca=\frac{U_{\mathrm{bubble}} \mu_{\mathrm{liq}}}{\gamma},
\end{equation}
where $U_{\mathrm{bubble}}$ is the propagation bubble velocity, $\mu_{\mathrm{liq}}$ is the
kinematic liquid viscosity and $\gamma$ is the interfacial surface tension between gas and liquid.

For example, the deposited film thickness
is proportional to $Ca^{2/3}$ in the range of small capillary numbers for circular channels
\cite{bretherton,heil-bretherton}. 
The problem of predicting flow patterns and associated mass transfers for the Bretheron-type flows
is of significant interests for chemical industry as it is widely used in chemical monolith
microreactors \cite{kreutzer-pressure-drop}. The large mass transfer can be achieved because of the
large interfacial area and small diffusion lengths \cite{cerro-bubble-train}. The heat transfer is
also increased in comparison with one phase flows \cite{fukugata-levelset}. While it
is possible to calculate the flow analytically for small capillary numbers \cite{bretherton}, it's
not possible to extrapolate it to the wider range of capillary and Reynolds numbers used in the
chemical
industry. Thus, the desire of consistent numerical simulations arises.

The two-dimensional geometry (circular tubes, parallel plates) flows are studied extensively in
experimental works of \citet{aussillous-deposition, cerro-bubble-train} and numerical works of
\citet{giavedoni-numerical,heil-bretherton}. All abovementioned works found that the Bretherton
analysis is valid only for small capillaries numbers $Ca\leq 0.003$ and deviates for larger ones.
That is caused by complex interplay of the gravitational,
interfacial, inertial and viscous forces \cite{gupta-review}. Historically, 
\citet{bretherton} indicated that inertia effects can be 
neglected. \citet{giavedoni-numerical} suggested
that
inertia effects
are negligible for $Ca \leq 0.05$ and have moderate changes for $Ca>0.05$ in the range of Reynolds
number from $0$ to $70$ for two-dimensional geometry flows. Later on, \citet{heil-bretherton}
extended
results for the flow between plates upto Reynolds number
$300$. The author indicates that while the influence of the established film thickness is
insignificant
($7$ percent from the film thickness measured at $Re=0$), the change in Reynolds number 
influences significantly the pressure distrubution and the flow field near the front bubble tip.
\citet{cerro-bubble-train} show that the mass transfer is strongly affected by flow direction for
small capillary numbers in case of upward and downward flows. Thus, to fully describe the
propagation of the semi-infinite finger in liquid media one needs to take into the account the
viscous, gravitational, surface and inertia forces \cite{gupta-review}.

In comparison with the two-dimensional Bretherton flow  (circular tubes, flow between
parallel
plates), there is a vast number of experimental
results available for the three-dimensional case as microchannels with square, triangular and
rectangular crosssections. For instance, \citet{cerro-bubble-train} performed a number of
experiments for a
bubble-train flow in capillaries of
circular and square cross section for horizontal, upward and downward flows.
\citet{shikazono-square} obtained
experimental
results for the deposition length dependency on the
capillary number for ethanol/air and water/air mixtures and for square, circular and triangular
shaped mirochannels. They also found experimental correlation for the prediction of bubble radii
based on the capillary number $Ca$ and the Weber number $We=Re\,Ca$.

In comparison with two-dimensional geometries, the experimental works
\cite{shikazono-square,cerro-bubble-train} supported by numerical simulations \cite{heil-threedim,
wang-non-circular} show interesting phenomena in three-dimensional geometries. It was found
\cite{heil-threedim,wong-films} that for rectangular
or square shaped capillaries there is a transition for a
certain capillary number, where the bubble crosssection changes from the non axisymmetric to axis
symmetric case. Non axisymmetric case is attributed to the case where the axial radius is
different from the
diagonal radius and the bubble has the non-circular shape in the channel crossection. In this case
the bubble shape mimics the shape of the square and looks like a rounded square. The dependence
of the diagonal and axial radii on the capillary number is shown in Fig. 
\ref{fig:heil:three:dim}.
\begin{figure}[ht]
\includegraphics[width=\textwidth]{Figures/capillary_width_heil.eps}
\caption{\citet{heil-threedim} results for the variation of the deposited liquid for the range of
capillary number. One can see the asymmetry between diagonal and axis diameter for the capillary
number $Ca\leq\hat{Ca}=0.04$. Courtesy of \citet{heil-threedim}. \label{fig:heil:three:dim}}
\end{figure}
The transition number between non-axisymmetric case and symmetric is indicated in a number of
works as $\widehat{Ca}=0.04$ \cite{cerro-bubble-train},
$\widehat{Ca}=0.1$
\cite{cerro-space,wang-non-circular}, $\widehat{Ca}=0.033$ \cite{heil-threedim}. If the capillary
number is larger
than
the critical capillary number, i.e. $Ca>\widehat{Ca}$, then the bubble becomes axisymmetric with the
radius of the droplet dependant on the capillary number. One of the example for the bubble radii
dependence on the capillary number is in Fig. \ref{fig:heil:three:dim}.

There are also a number of numerical works for the three-dimensional case. For instance, 
\citet{wong-films,wong-pressure} studied
three-dimensional bubbles in
polygonal capillaries and calculated bubble shapes for different
slug and bubble cross sections and menisci appearance.
\citet{heil-threedim} performed three-dimensional simulations for circular shaped,
square and rectangular shaped capillaries. They indicated the transition
capillary number where the streamlines pattern
changes from having a vortex in front of the bubble and not having it for larger capilllary
numbers, {\color{red} Fig. \ref{fig:streamlines:pattern} }. For the square channel the critical
capillary number is $Ca=0.691$.  Authors also found an empirical correlation which
allows to put
different aspect ratio rectangular channels simulation results on one
curve. It is also indicated that for certain aspect ratio microchannels $\alpha=\frac{a}{b}\geq
2.04$ the interface can not make up the
axisymmetric case for any capillary number. \citet{wang-non-circular} performed numerical
simulations by the VOF technique for the
non-circular
shaped capillaries as square and triangular shaped capillaries. They also investigated the relative
velocity of the bubble for
different capillary numbers.

As it is mentioned by \citet{gupta-review}, the most common techniques to simulate the Bretherton
phenomena are the Volume of Fluid (VOF) method \cite{wang-non-circular}, the level-set method
\cite{fukugata-levelset} and the finite element methods \cite{kreutzer-taylor,heil-threedim}. It is
also indicated \cite{gupta-review}, that the new techniques which are still in the development stage
are the lattice Boltzmann method and phase field methods \cite{anderson-diffuse,gurtin-binary}.
While the finite element methods solve the Bretherton flow as a free-surface problem with a sharp
interface but without underlying details for gas, the lattice Boltzmann method which is a continuous
interface method arises as a promising alternative for simulation of gas finger propagation. The
continuous interface has more flexebility in simulations involving coalescence and/or droplet
breakup.

In our previous work \cite{kuzmin-binary2d} we already covered the Bretherton flow between plates.
It was shown that the free-energy binary liquid model, which is a phase field method, simulates
with reasonable accuracy the Bretherton/Taylor phenomenon. The goal of this work is to examine the
three-dimensional geometry simulations conducted by the free-energy binary liquid lattice Boltzmann
method. We focus on the feasibility study of the numerical method for the microchannels with
square crosssections.

The lattice Boltzmann method has emerged as a successful method to simulate
a wide variety of phenomena including hydrodynamics \cite{yu}, thermal flows
\cite{karlin-minimalmodels}, microflows \cite{ansumali-small-knudsen},
ferrofluids \cite{kuzmin-aniso}, and multiphase flows
\cite{swift,Shan-chen:extended}. The method is a particle method which allows to simulate physical
phenomena on the microscopic level. For instance, the introduction of the force or potential on the
microscopic level allows to restore multiphase macroscopic equations \cite{swift,
Shan-chen:extended}.

The binary liquid free-energy LB model due to \citet{swift} we used
simulates two liquids with the assumption of uniform overall
density. While the classical Bretherton problem is stated for gas and liquid, which are of
significantly
different densities and viscosities, the parameters were carefully chosen as to neglect the inertia
effects, see Section \ref{sec:numerical:benchmark}. Moreover, the results of
\citet{giavedoni-numerical} and \citet{heil-bretherton} show
negligible Reynolds number effects on the film thickness for a relatively wide range of Reynolds
numbers. Thus, a major governing factor for microchannel flows is not the density ratio, but
the
viscosity ratio. That is why the uniform density binary liquid model is suitable for this kind of
simulations. The goal of this work is to do a feasibility study of the LBM
binary-liquid model to simulate and correctly predict flow patterns for the Bretherton/Taylor
problem. The work results are in good agreement with other simulations
\cite{heil-threedim,wang-non-circular}.

One should also acknowledge the works of \citet{pagonabarraga-fingers} on menisci
in thin films for fingering phenomena. \citet{sehgal-microchannel} performed lattice Boltzmann
simulations of two-dimensional channel flows for relatively large capillary numbers, and
found discrepancies with the classical Bretherton theory, which
is limited to the low capillary number regime \cite{giavedoni-numerical}. The Shan-Chen model was
used to simulate the Bretherton problem, which is sometimes referred to contain thermodynamically
inconsistent interface \cite{nourgaliev-breakup}.

The paper is organized as follows.  First, we explain the simulation benchmark construction
mainly based on our previous work \cite{kuzmin-binary2d}. Then, the binary liquid lattice
Boltzmann model is outlined. The simulation results for the three-dimensional case are presented in
the results section. The results cover the film thickness dependency on the capillary number, the
relative velocity of the bubble and the vortex profiles. The paper
is concluded with a summary of the main findings.
%The critical capillary numbers for changing velocity pattern depending on the eccentricity
%parameter $\alpha$ are presented in Table \ref{table:recirculation:data}.
%\begin{table}
%\begin{tabular}{c|c}
%$\alpha$& $Ca_{T}$\\
%\hline
%1& 0.691\\
%1.1 & 0.688\\
%1.3 & 0.666\\
%1.5 & 0.631\\
%\end{tabular}
%\caption{The results for the recirculation region.\label{table:recirculation:data}}
%\end{table}

\section{Numerical benchmark approach}
\label{sec:numerical:benchmark}
The main discussion here follows and is based on the two-dimensional benchmark approach
\cite{kuzmin-binary2d}. It was indicated that the benchmark layout should have certain
dimensions to conduct simulations. The classical Bretherton benchmark layout is represented in Fig.
\ref{fig:classical:benchmark}. It describes the gas finger propagation through the liquid media.
The film thickness in this case is measured at the inlet. In lattice Boltzmann framework such
formulation has certain challenges \cite{kuzmin-binary2d}. Some of them are attributed to the
dynamic coupling of the inlet and outlet conditions \cite{giavedoni-numerical}. Instead, we
proposed the numerical benchmark indicated in Fig. \ref{fig:lbm:benchmark}. The dimensions of the
channel are chosen as $H_{\mathrm{eff}}\times H_{\mathrm{eff}} \times 15 H_{\mathrm{eff}}$. The
initialized bubble length is taken as $5 H_{\mathrm{eff}}$. \citet{giavedoni-numerical} showed that
the film stabilizes at the distances of $2.6-4.0$ diameters
from the front tip depending on the Reynolds number. \citet{heil-threedim} measure the bubble radii
at the distance as $5.5 H_{\mathrm{eff}}$ and indicate it to be sufficient for $Ca\leq 10$. In
this work the conducted simulations capillary number is not larger than unity $Ca\leq 6$.  Thus, the
film thickness is chosen to be measured in the
middle of the bubble, which is located at least at a distance of $2.5-3$ channel heights from the 
bubble tip.%at the distances one channel height from the rear tip upto one channel

For simplicity, the periodic boundary conditions are applied. Thus, the bubble train is simulated.
To avoid the mutual influence of bubbles on each other the channel length was chosen as $15
H_{\mathrm{eff}}$. Periodic boundary conditions imply as well that the fluid is accelerated by a
body force in the framework of the LBM but not with the pressure difference. Therefore, we can not
address simulations with upward and downward flows \cite{cerro-bubble-train}. It is assumed that
the gravity effects are negligible and the flow is in the horizontal direction driven by the body
force. 
\begin{figure}[ht]
\includegraphics*[bb=153 610 405 717,width=\textwidth]{Figures/benchmark_classical.eps} 
\caption{The classical benchmark layout is presented . The semiinfinite gas bubble
propagates through the liquid media. The inlet pressure is specified $P_{in}$, the outlet is the
free outflow. The bubble propagates with constant velocity $U_{\mathrm{bubble}}$. 
\label{fig:classical:benchmark}}
\end{figure}
\begin{figure}[ht]
\includegraphics*[bb=152 470 410 713,width=\textwidth]{Figures/benchmark_lbm.eps}
\caption{The lattice Boltzmann benchmark layout is used in present calculcations. The dimensions of
the domain are chosen as $H_{\mathrm{eff}}\times H_{\mathrm{eff}}\times 15 H_{\mathrm{eff}}$. The
length of the bubble was chosen as $5 H_{\mathrm{eff}}$ for the film thickness to stabilize.
Periodic boundary conditions are applied in the streamwise direction. The flow is driven by body
force. \label{fig:lbm:benchmark}}
\end{figure}

It was indicated \cite{kuzmin-binary2d} that the film thickness for two-dimensional simulations should be at least as twice large
as the interface thickness. The interface thickness in the continuous formulation of the
free-energy binary liquid formulation occupies $3-5$ nodes. Therefore, the film should be
resolved at least as $6-10$ nodes.

\cite{heil-threedim} specify that the diagonal bubble radius is different from the
axial bubble radius for $Ca\leq \widehat{Ca}\approx 0.05$. The radii coincide for larger values of
$Ca$, but the value of bubble radii is $R_{diag}=R_{axis}=0.99 H_{\mathrm{eff}}$. Therefore, taking the minimal requirement for the film thickness to be
resolved, i.e. $6$ lattice nodes, one can obtain the grid size as $600\times 600 \times 9000 $ to
properly resolve the binary liquid microchannels bubbles motion. This size implies highly
computational resource to be available for simulations.

One can argue that the non-uniform grid can be used. However, the deposited film thickness is the
function of the curvatures of the front bubble tip \cite{bretherton}. Therefore, the grid needs to
be refined just at the interface between gas and liquid. It is complicated to achieve for the
dynamic systems where a bubble moves in the streamwise direction.

Also, the binary liquid model is used. The model has an uniform density. A part of discussion is
addressed in the introduction that to fully address the Bretherton problem one needs to take into
the account gravitational, viscous, inertial, and surface tension forces. However, 
\citet{shikazono-square} indicated that for square shaped capillaries radii have the
following dependency on the capillary number $Ca$ and the Weber number $We$:
\begin{equation}
\label{radii:experimental}
\begin{aligned}
&R_{diag}=1.171-\frac{2.43 Ca^{2/3}}{1+7.28 Ca^{2/3}-0.255 We^{0.215}}\\
&R_{axis}=
\begin{cases}
1, &R_{diag}>1\\
R_{diag}, &R_{diag}\leq 1.
\end{cases}
\end{aligned}
\end{equation}
Equation (\ref{radii:experimental}) shows that a change in the Reynolds number $0\leq Re \leq 100$
implies the change of the radii from $R_{diag}=R_{axis}=0.96$ to $0.93$ for $Ca=0.1$ and from
$R_{diag}=R_{axis}=0.88$ to $0.85$ for $Ca=0.1$. The error of $3$ percent of the channel height is
acceptable for current simulations where the Reynolds number is less than $10$.

Thus, having limitations of the computational memory and performance requirements the focus of the
present work is a feasibility study of the binary liquid lattice Boltzmann method for the
microchannel simulations with square crossections conducted in the moderate capillary number range
$0.1 \leq Ca \leq 1.0$. We base our work on comparison with the already available studies.

To further reduce the computational overhead only quarter of the channel is simulated, see Appendix
\ref{append:sym}. The next section briefly explains the binary liquid lattice Boltzmann model. Then
the steady-state approach, the dependency of the radii on the moderate capillary number values
$0.1\leq Ca \leq 1.0$, the velocity pattern, and the relative bubble velocity, see Section
\ref{sec:results}.

\section{Binary liquid lattice Boltzmann model}
The lattice Boltzmann equation (LBE) operates on a rectangular grid representing the
physical domain. It utilizes
probability distribution functions (also known as particle populations)
containing information about
macroscopic variables, such as fluid density and momentum. LBE consists of
two parts: a local collision step, and a propagation step which transports
information from one node to another in certain
directions specified by the discrete velocity set.
The LBE is typically implemented as follows \cite{ginzburg-boundary-main}:
\begin{equation}
\label{standard:implementation}
\begin{aligned}
&f_i^{*}(\bm{x},t)=\omega f_i^{eq}(\bm{x},t)-(1-\omega) f_i(\bm{x},t) +
F_i,&&\text{collision step}\\
&f_i(\bm{x}+\bm{c_i},t+1)=f_i^{*}(\bm{x},t),&&\text{propagation step}\\
&g_i^{*}(\bm{x},t)=\omega_{\phi} g_i^{eq}(\bm{x},t)-(1-\omega_{\phi})
g_i(\bm{x},t),&&\text{collision step}\\
&g_i(\bm{x}+\bm{c_i},t+1)=g_i^{*}(\bm{x},t),&&\text{propagation step},
\end{aligned}
\end{equation}
where $\{f,g\}_i$ are the probability distribution functions in the direction $\bm{c_i}$, $\omega$
is the
relaxation parameter, $\omega_{\phi}$ is the phase relaxation parameter, and $F_i$ is the external
force population which
mimics the force behavior, Eq. \ref{full:navier:stokes}.

The binary fluid LB model is
based on a free-energy functional \cite{swift,landau}, and operates with two
sets of populations: one to track the pressure and the velocity fields, and another to represent the
phase field $\phi$ indicating gas or liquid.
The equilibrium populations \cite{pooley-contact} are defined as:
\begin{equation}
\label{set:equilibrium:binary}
\begin{aligned}
&f_i^{eq}&&=w_i 
\biggl(3
p_0 - k \phi \Delta \phi
+\frac{u_{\alpha}c_{i\alpha}}{c_s^2}+\frac{Q_{i\alpha\beta}u_{\alpha } u_ {
\beta}}{2 c_s^4}\biggr)\\
&&&+k w_i^{\alpha\beta} \partial_{\alpha} \phi\partial_{\beta} \phi, 1\leq i \leq Q-1\\
&f_0^{eq}&&=\rho-\sum_{i\neq0}{f_i^{eq}}\\
&g_i^{eq}&&=w_i\left(\Gamma \mu + \frac{\phi c_{i\alpha} u_{i\alpha}}{c_s^2}+\phi
\frac{Q_{i\alpha\beta}u_{\alpha}u_{\beta}}{2 c_s^4}\right), \leq i \leq Q-1\\
&g_0^{eq}&&=\phi-\sum_{i\neq0}{g_i^{eq}}\quad,
\end{aligned}
\end{equation}
where $\Gamma$ is the mobility parameter; the chemical potential
$\mu=-A\phi+A\phi^3-k\Delta\phi$; $k$ is the parameter related to the surface
tension; $A$ is the parameter of the free-energy model; $Q$ is the number of the directions ($19$
for the $D3Q19$ model); the tensor
$Q_{i\alpha\beta}=c_{i\alpha} c_{i\beta} - c_s^2 \delta_{\alpha\beta}$ with
the sound speed parameter $c_s^2=1/3$. The bulk pressure
is expressed as $p_0=c_s^2 \rho +A (-0.5 \phi^2+0.75 \phi^4)$. Velocity set is defined as:
\begin{equation} 
\begin{aligned}
&c_{ix}=\{0,1,-1,0, 0,0, 0,1,-1, 1,-1,0, 0, 0, 0,1,-1, 1,-1\}\\
&c_{iy}=\{0,0, 0,1,-1,0, 0,1, 1,-1,-1,1,-1, 1,-1,0, 0, 0, 0\}\\
&c_{iz}=\{0,0, 0,0, 0,1,-1,0, 0, 0, 0,1, 1,-1,-1,1, 1,-1,-1\}.
\end{aligned}
\end{equation}
The weights are $w_0=0$, $w_{1-6}=\frac{1}{6}$ and $w_{7-18}=\frac{1}{12}$. The weights
related to the inclusion of the surface tension are:
\begin{equation}
\begin{aligned}
&w^{xx}_{1-2}=w^{yy}_{3-4}=w^{zz}_{5-6}=\frac{5}{12}\\
&w^{xx}_{3-6}=w^{yy}_{1-2,5-6}=w^{zz}_{1-4}=-\frac{1}{3}\\
&w^{xx}_{7-10,15-18}=w^{yy}_{7-14}=w^{zz}_{11-18}=-\frac{1}{24}\\
&w^{xx}_{11-15,}=w^{yy}_{15-18}=w^{zz}_{7-10}=\frac{1}{12}\\
&w^{xy}_{7,10}=-w^{xy}_{8,9}=w^{yz}_{11,14}=-w^{yz}_{12,13}=w^{zx}_{15,18}=-w^{zx}_{16,17}=\frac{1}{
4}.
\end{aligned}
\end{equation}

The set of equations (\ref{set:equilibrium:binary}) restores the macroscopic
fluid equations as:
\begin{equation}
\label{full:navier:stokes}
\begin{aligned}
&\partial_t \rho+ \partial_{\alpha} \rho u_{\alpha}=0\\
&\rho\left(\partial_t+u_{\beta}\partial_{\beta}\right) u_{\alpha}=F_{\alpha}
-\partial_{\beta}P_{\alpha \beta} +
\nu\partial_{\beta}\left(\rho\partial_{\alpha}u_{\beta}+\rho\partial_{\beta} u_{\alpha}\right)\\
&\partial_t \phi + \partial_{\alpha} \phi u_{\alpha}=M \partial^2_{\beta\beta} \mu,
\end{aligned}
\end{equation}
where $\nu=c_s^2 (\tau-1/2)$ is the dynamic viscosity,
$M=\Gamma(\tau_{\phi}-1/2)$ is the mobility parameter, and $\tau=\frac{1}{\omega}$ and $\tau_{\phi}$
are the relaxation parameters of density and phase fields. The first equation of system
(\ref{standard:implementation}) simulates the continuity equation and the Navier-Stokes equation.
The lattice Boltzmann update rule for the second set $g_i$ simulates the phase propagation with the
supplied velocity from the Navier-Stokes equation.

The system allows the separation of the liquid
phase with $\phi=1$ and a so-called gas phase with $\phi=-1$. The
relaxation time is taken as linearly dependent on the relaxation
times $\tau_{\mathrm{gas}}$ and $\tau_{\mathrm{liq}}$:
$\tau=\tau_{\mathrm{gas}}+\frac{\phi+1}{2}(\tau_{\mathrm{liq}}-\tau_{\mathrm{gas}})$. This allows
to change viscosity from the gas viscosity
$\nu_{\mathrm{gas}}=\frac{1}{3}\Bigl(\tau_{\mathrm{gas}}-\frac{1}{2}\Bigr)$ to the liquid viscosity
$\nu_{\mathrm{liq}}=\frac{1}{3}\Bigl(\tau_{\mathrm{liq}}-\frac{1}{2}\Bigr)$ while phase changes
accordingly.The surface tension in the framework of the binary liquid model is $\sqrt{\frac{8 k
A}{9}}$.

\section{Results}
\label{sec:results}
This section describes numerical simulations. One can refer to the appendix \ref{append:scaling} for
initialization and scaling procedures. The simulations were conducted on different grids
corresponding to fluid domain grids as $100 \times 100 \times 1500$, $160 \times 160 \times 1500$,
$160 \times 160 \times 2400$ and $200 \times 200 \times 3000$. The body forces were varied as
$\frac{\mathrm{d}P}{\mathrm{d}z}=10^{-6}-10^{-4}$. The binary-liquid parameters were kept to be in
the stable region $k=A=0.004,0.04$. The relaxation rates were taken as $\tau_{\mathrm{liq}}=2.5$
and $\tau_{\mathrm{gas}}=0.7$ giving the gas-liquid dynamic viscosity ratio $10$. The relaxation
parameter for the phase field was $\tau_{\phi}=1.0$.

First, we examine when the steady-state approach is
achieved. Then, the critical capillary number is identified when the transition from the
non-symmetrical to axisymmetrical bubble shape occurs. The dependency of the radii on the capillary
number is presented for the range of moderate capillary numbers $0.1 \leq Ca 1.0$. Results are
concluded with studies of the velocity pattern.

\subsection{Steady-state approach}
\label{sec:steady:state}
We performed different simulations to understand a number of steps required for the problem to come
on the steady-state. The grid to be simulated is $52\mathrm{x}52\mathrm{x}750$ which simulates the
quarter of the channel with the initial
width as $6.5$ lattice Boltzmann units together with the body force as
$\frac{\mathrm{d}P}{\mathrm{d}x}=1.6
\mathrm{x}10^{-6}$. The simulation was performed for $140000$, $160000$, $180000$, $200000$,
$220000$ and
$240000$ iterations. The results are summarised in Table \ref{table:steady:state}. Whole the
capillary number varies only in the 4th digit, radiuses are one percent accurate. 
\begin{table}
\begin{tabularx}{\textwidth}{|X|X|X|X|}
\hline
$N_{iter}$&$Ca$&$R_{axis}$&$R_{diag}$\\
\hline
$140000$&$0.1879$&$0.8657$&$0.8708$\\
$160000$&$0.1878$&$0.8650$&$0.8702$\\
$180000$&$0.1882$&$0.8647$&$0.8700$\\
$200000$&$0.1893$&$0.8645$&$0.8698$\\
$220000$&$0.1925$&$0.8644$&$0.8697$\\
$240000$&$0.1961$&$0.8645$&$0.8697$\\
\hline
\end{tabularx}
\caption{Results for the steady-state case. One
can see that $200000$ steps are enough to structure
the bubble and approach the steady state. The small noise in the capillary number is connected to
the identification of the interface. One needs to address the interpolation and the spurious
currents existing in the system. All other parameters are relaxed to the
steady-state.\label{table:steady:state}}
\end{table}

To further examine the steady-state the velocity in the streamwise direction is plotted. The values
of the velocity are taken on the contour where phase values are taken as $\phi=0$. This corresponds
 to the bubble interface. One can see the contour plot and the corressponing streamwise velocity
component in Fig. \ref{fig:velocity:contour}. The crossection is a plane $x=0$, where $x$ points
towards the reader, $z$ is the streamwise direction. As far as fluxes inside the bubble has
clockwise and counterclockwise directions, see Section \ref{sec:velocity}, one needs to thoroughly
examine only the points corresponding to the front and the rear bubble tips. If one finds it equal
to each other, then it's exactly the velocity with which the bubble propagates in the microchannel.
For physical analogy, one can imagine the rotating ball propagating in the streamwise direction. 
\begin{figure}[ht]
\includegraphics[width=\textwidth]{Figures/velocity_interface_contour.eps}\\
\includegraphics[width=\textwidth]{Figures/velocity_interface_values.eps}\\
\caption{The bubble shape and the corresponding streamwise velocities taken on the bubble
interface after $240000$ iterations. Directions $y$ and $z$ are scaled to the half of the channel
$H_{\mathrm{eff}}/2$. One can see that the bubble front and rear tips are propagating with the
same velocities, which is actually the bubble velocity. For more details on velocity patterns
see Section \ref{sec:velocity}. \label{fig:velocity:contour}}
\end{figure}

\subsection{Radii transition}
Due to performance issues we can only access the
transition region, i.e. $Ca\geq 0.05$. In this region the proper resolution of the film thickness
is guaranteed. To identify the critical capillary number $\widehat{Ca}$ a number of simulations
were conducted. Surprisingly the transition occurs in the region computationally accessible for
current simulations. In comparison with the results of the \citet{heil-threedim}
($\widehat{Ca}=0.04$) our
results are more close to the VOF continuous interface method simulation by
\citet{wang-non-circular} ($\widehat{Ca}=0.1$) and experimental dependency of
\citet{shikazono-square}, see Table
\ref{table:transition:results}. The transition happens for $Ca\approx 0.09$.  One can see
two examples of non-symmetric and axisymmetric bubble shapes for $Ca=0.053$ and $Ca=1.13$, see Fig.
\ref{fig:crosssections:sym}.  
\begin{table}
\begin{tabularx}{\textwidth}{|X|X|X|X|X|X|}
\hline
$Ca$&$R_{axis}$&$R_{diag}$&$Re$&$R_{axis,Wang}$&$R_{diag,Wang}$\\
\hline
$0.053$&$0.95$&$1.01$&$0.907$&$0.99$&$0.99$\\
$0.078$&$0.93$&$0.95$&$1.062$&$0.96$&$0.96$\\
$0.132$&$0.89$&$0.89$&$1.80$&$0.93$&$0.93$\\
\hline
\end{tabularx}
\caption{Simulation results in terms of $R_{axis}$ and $R_{diag}$ for the transition region between
non-symmetric and axisymmetric cases. $R_{axis,Wang}$ and $R_{diag,Wang}$ are experimental
correlations by \citet{wang-non-circular}. \label{table:transition:results}}
\end{table}
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/phase_crossection_ca5.eps}
\includegraphics[width=0.5\textwidth]{Figures/phase_crossection_ca13.eps}\\
\caption{Crosssections of the phase field in the middle of the bubble for $Ca=0.053$
(left) and $Ca=1.13$ (right). Crosssections are rescaled on the half channel height. Other
parameters are indicated in Table \ref{table:transition:results}. One can see that the left picture
is asymmetric in comparison
with the right one. The transition happens for $Ca\approx 0.09$.\label{fig:crosssections:sym}}
\end{figure}

\section{Velocity pattern}
\label{sec:velocity}
The velocity pattern is known to change its behavior depending on the capillary number. For
relatively low capillary number $Ca<0.6$ there exists a vortex in front of the bubble in the moving
frame moving with the bubble \cite{heil-threedim}. For larger ones it is indicated that there is no
vortex in front of the bubble. The transition capillary number is indicated by
\citet{heil-threedim} and it equals for the rectangular channel $Ca=0.691$. However, the
distribution of vorticity strongly depends on Reynolds number \cite{heil-bretherton}, as well the
pressure distribution. In the case of the bubble train it is indicated by \citet{kreutzer-taylor}
that the pressure is largely influenced by bubble frequency and slug distance. The present
simulations are conducted for bubble trains and certain differences in terms of changing streamline
patterns are expected. We examined velocity patterns to identify the moment of the streamlines
pattern change. We chose two
representative capillary numbers as $Ca=0.47$ and $Ca=0.63$. 
One can
see a clear transition between associated patterns. The transition capillary number
\cite{heil-threedim} is slightly higher than the simulation results. In present simulations the
transition occurs at $Ca=0.5-0.6$.
\begin{figure}[ht]
\includegraphics[angle=90,width=\textwidth]{Figures/stream_ca47.eps}\\
\includegraphics[angle=90,width=\textwidth]{Figures/stream_ca63.eps}\\
\caption{Velocity vector maps for $Ca=0.47$ and $Ca=0.63$. One can see that there are no vortexes
created
before the bubble for $Ca=0.63$. The transition happens between $Ca=0.47$ and $Ca=0.63$, which
is a different value from $Ca=0.69$ \cite{heil-threedim}. However, we performed a number of
different types of simulations (different force, width initialization). The critical transition
capillary number $\widehat{Ca}$ is not well imposed and varies from $\widehat{Ca}=0.6-0.7$.
\label{fig:streamlines:pattern}}
\end{figure}

\section{Variation bubble length}
The work of \citet{wang-non-circular} shows the variation of the bubble
thickness through the length of
bubble. The bubble shape in terms of axial and diagonal radii is reconstructed for a number of
capillary numbers, Fig. \ref{fig:bubble:variation:capillaries}. One can see that for smaller
capillary numbers the diagonal radius exhibits a small jump near the rear bubble tip. It coincides
with results of \citet{wang-non-circular}. However, present simulations show smoother behavior of
the jump.  
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/bubble_rad_axis.eps}
\includegraphics[width=0.5\textwidth]{Figures/bubble_rad_diag.eps}\\
\caption{Radii variations along the bubble for different capillary numbers in the plane $x=0$.
One can see that the diagonal radius shows the jump near the bubble rear tip. It coincides with the
VOF simulations of \citet{wang-non-circular}.Coordinate $z$ increases in the streamwise direction.
\label{fig:bubble:variation:capillaries}}
\end{figure}

Another interesting phenomenon for the bubble shape was indicated by \citet{heil-threedim}. In
their simulations the non-symmetric shape was observed even for larger capillary number $Ca>4$ near
the bubble front tip which after certain distance decay to have radii equal to each other. The
bubble shape was calculated for the large capillary number $Ca=6.43$, see Fig.
\ref{fig:bubble:ca:large}. However, even the diagonal radius shows the jump near the front tip of
the bubble and it resembles the shape given by \citet{heil-threedim}, but the difference is around
$1$ percent and can be attributed to the linear interpolation error used for the interface
calculation. 
\begin{figure}[ht]
\includegraphics[width=\textwidth]{Figures/bubble_ca_large.eps}
\caption{The bubble shape for the capillary number $Ca=6.43$ is presented.$z$ increases in the
streamwise direction. One can see that near the front tip (large $z$) the diagonal radius exhibits
a jump which is relaxed after to the axis radius. Even this resembles
the shape of the curves obtained by \citet{heil-threedim}, but the
difference between radii is around $1$ percent and can be attributed to
the linear interpolation used in the calculation of the interface.\label{fig:bubble:ca:large}}
\end{figure}

\section{Capillary number}
The purpose of this section is to compare radii dependence on the capillary number. The film
thicknesses were extracted from the middle of the bubble. The data presented in Fig.
\ref{fig:capillary:comparison} is simulated with different techniques including different
initialization, different grids as $52 \times 52 \times 1500$, $82 \times 82 \times 1500$ and $82
\times 82 \time 2400$. All the results are consistent. Therefore for the capillary number $Ca\geq
0.05$ the grid independence is achieved. The results are compared with results of
\citet{heil-threedim} and of \citet{wang-non-circular}. The data was extracted from the
correspondent references with the help of the program ``Engauge Digitizer''. The lattice Boltzmann
can catch the transition between non-symmetric to axisymmetric shape of the bubble.

Present simulations show that the corresponding critical capillary number is close to
$\widehat{Ca}=0.1$ which is close to work of \citet{wang-non-circular}. 
In comparison with
the continuous method VOF \cite{wang-non-circular} present simulations show less noisy trend and
more accurate transition.

For moderate capillary number $Ca\geq 0.2$ radii show less value than the literature results. That
might be corresponding to the bubble train simulation or the influence of the binary-liquid model. 
\begin{figure}[ht]
\includegraphics[width=0.5\textwidth]{Figures/capillaries_comparison_heil.eps}
\includegraphics[width=0.5\textwidth]{Figures/capillaries_comparison_wang.eps}\\
\caption{The comparison for the axial and the diagonal radii
versus capillary numbers between present simulations and the results of \citet{heil-threedim}
(top), of \citet{wang-non-circular} (bottom). One can see that
the code mimics
behavior of the earlier published results.  \label{fig:capillary:comparison}} 
\end{figure}

\section{Relative velocity}
One of the differencies between the flow in tubes and square cappillaries is that the flow has
different patterns. In square microchannels the main mass flow goes through corners
\cite{heil-threedim,wang-non-circular}. This affects the mass flow significantly. One of the
characteristics of the mass flow is the normalized relative velocity dependence on the capillary
number. Bubble moves always faster than surrounding liquid. The following quantity is defined as
relative velocity \cite{cerro-bubble-train}:
\begin{equation}
W=\frac{U_{\mathrm{bubble}}-U_{ls}}{U_{\mathrm{bubble}}},
\end{equation}
where $U_{\mathrm{bubble}}$ is the bubble interface velocity, $U_{ls}$ is the liquid superficial
velocity. As far as the steady state is achieved one can take the liquid superficial velocity
exactly at the distance in the middle between bubbles. Then, the liquid superficial velocity is
defined as:
\begin{equation}
U_{ls}=\frac{\int_{A}{U_{liq}dA}}{A}.
\end{equation}
The relative velocity was calculated and compared with the results of \citet{wang-non-circular},
see Fig. \ref{fig:relative:velocity}. One can see the agreement between simulations indicating that
the mass transfer coefficients can be captured accurately. 
\begin{figure}[ht]
\includegraphics[width=\textwidth]{Figures/relative_velocity.eps}
\caption{Relative velocity comparison between current simulations and simulations of
\citet{wang-non-circular}. One can see a qualitative agreement. \label{fig:relative:velocity}}
\end{figure}

The relative velocity is very important quantity in the mass flow characterization
\cite{kreutzer-taylor,yue-mass}. The frequency of the bubbles \cite{kreutzer-taylor} and the
inertia effects \citet{heil-bretherton} influence significantly the mass transfer coefficient. The
scope of this work is the feasibility study of the lattice Boltzmann model to be applied for
three-dimensional microchannels with square crosssections. The future work is planned to study mass
transfer coefficient and solve associated advection-diffusion coefficient. However, the preliminary
studies of the relative velocity show that the binary liquid model can be used for accurate mass
transfer characterization.

% \section{Train simulation}
% \label{section:bubble:train}
% As far as the requirements for the grid are ``heavy'' we performed a bunch of simulations to
% examine how the distance between bubbles influences on the thickness. For this purpose the
% simulation was conducted on the grid 82x82x750. We examined different initialized distances
%between
% bubbles. The initial lengths of bubbles were chosen as $300$, $350$, $400$, $450$, $500$, $550$
% lattice units. Therefore, the corresponding initial distances between bubbles are
% $450$,$400$,$350$,$300$,$250$ lattice units. On the steady state regime the distances become
%$386$,
% $317$, $248$,$176$,$99$,$23$ lattice units. The corresponding capillary numbers calculated based
%on
% the center bubble velocity are as follows $0.905$, $0.986$,$1.08$,$1.19$, $1.33$, $1.49$. The
% calculated radiuses (diagonal radius equal to the axes radius) are as follows
% $0.783$,$0.778$,$0.772$,$0.764$,$0.753$,$0.747$. Thus, one can see the great impact of the bubbles
% on each other velocity for the given body force. However, in terms of the film thicknesses those
% numbers look adequate. One can see the film thicknesses dependency on the capillary number for
% simulations obtained by \citet{heil-threedim} and bubble train simulations data in Fig.
% \ref{fig:capillaries:train} and in Fig. \ref{fig:capillary:comparison}. One can see that the
% results are consistent and in a good agreement.
% Therefore, though the bubble train mutual motion influences on the film thickness by increasing
%the
% associated capillary number,  however, the film thickness is in accordance with the capillary
% number correlation. Thus, the decrease of the numerical domain is quantified and can be used to
% obtain the data. However, one should be cautious about the initialization, because the Poiseuille
% flow initialization is deviated even further due to mutual motion of bubbles.
% \begin{figure}
% \includegraphics[width=0.97\textwidth]{Figures/capillaries_comparison_train.eps}
% \caption{The comparison for the film thickness between \citet{heil-threedim} data and bubble train
% simulations. Though, one can see that the decrease of the distance between bubbles causes the
% increase in the associated capillary number, however, the correlation dependance between film
% thickness and capillary number stays the same. This can help for simulations to reduce the
% numerical domain. \label{fig:capillaries:train}}
% \end{figure}

\section{Conclusion}
The present work conducts the binary-liquid simulations for three-dimensional channels with square
crosssections with the lattice Boltzman method. The
method is a continuous interface method. By resolving properly the film thickness as twice as the
interface thickness \cite{kuzmin-binary2d} the results were shown to be in consistence with the
available in literature results. The film thickness dependency on the capillary number, the
transition from the axisymmetric to symmetric case, the velocity pattern were investigated. While
the goal of this work is the feasibility study, the more numerical studies are needed to
understand the influence of the slug length, inertia effects, pressure distributions
\cite{kreutzer-taylor,yue-mass} to design the microchannels. Results evedince that the lattice
Boltzmann binary liquid model can be used for simulation of gas bubbles in microchannels.

\appendix
\section{Scaling procedure}
\label{append:scaling}
The thorough scaling procedure is described in the work \cite{kuzmin-binary2d}. Here only the
outline is presented on how the simulations are initialized:
\begin{description}
 \item[Capillary number] One first needs to set the capillary number and the Reynolds number for
simulations. 
 \item[Film thickness] After the capillary number is prescribed, one needs to approximate the axial
bubble radius $R_{axial}$, the corresponding film thickness is $\delta=1-R_{axial}$. One can use
correlations of \citet{shikazono-square} or of \citet{kreutzer-taylor}. It can be done either by
taking correlations from numerical simulations by
\citet{heil-bretherton} and \citet{giavedoni-numerical}.  
\item[Grid choice] After specifying the film thickness, one needs to choose the associate
number of nodes to resolve the film thickness. For binary liquid parameters used in simulations,
$A=0.004,0.04$ and $k=0.004,0.04$ the
interface is spread over approximately $5$ lattice units. To obtain grid independent results one
needs to choose the
film thickness to be $2-2.5$ times larger than the interface thickness \cite{kuzmin-binary2d}.
Giving the number of nodes to resolve the film thickness and giving the lattice Boltzmann
benchmark, see Section \ref{sec:numerical:benchmark}, one can obtain the grid dimensions.

\item[Velocity] The Reynolds and the capillary numbers are supplied from the physical world. We
accept the following definitions:
\begin{equation}
\label{definitions:numbers}
\begin{aligned}
&Ca=\frac{U_{\mathrm{bubble}} \mu_{\mathrm{liq}}}{\gamma} \\
&Re=\frac{U_{\mathrm{bubble}} H_{\mathrm{eff}}}{\mu_{\mathrm{liq}}}.
\end{aligned}
\end{equation}
In Eq. \eqref{definitions:numbers} the interface tension coefficient is prescribed by parameters of
the binary liquid model $k$ and $A$, the effective channel height $H_{\mathrm{eff}}$ is obtained
from the film thickness resolution, $\mu_{\mathrm{liq}}$ is defined through the relaxation rate
$\tau_{\mathrm{liq}}$. The relaxation rate $\tau_{liq}$ is prescribed from the gas-liquid viscosity
ratio $\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}$, where $\mu_{\mathrm{gas}}$ is defined through
the gas relaxation rate $\tau_{\mathrm{gas}}$. Parameter $\tau_{\mathrm{gas}}$ have constraints
$\tau_{gas}\geq 0.5$, therefore constraining the linked parameter $\tau_{\mathrm{liq}}$. Moreover,
the stability of the lattice Boltzmann scheme is known to deteriorate with the relaxation parameter
$\tau_{\mathrm{liq},\mathrm{gas}}\approx 0.5$ \cite{kuzmin-trt-stability}. On the other hand the
accuracy of the lattice Boltzmann liquid model is detiorated as
$\Bigl(\tau_{\mathrm{liq},\mathrm{gas}}-\frac{1}{2}\Bigr)^2$ \cite{ginzburg-trt-simple-hydro}. Thus,
there is a compromise of the choice of parameters $\tau_{\mathrm{liq}}$ and $\tau_{\mathrm{gas}}$.
The gas-liquid viscosity ratio in the current simulations was chosen as
$\frac{\mu_{\mathrm{liq}}}{\mu_{\mathrm{gas}}}=10$ and the relaxation parameters were chosen to be
in the safe region, i.e. $\tau_{liq}=2.5$ and $\tau_{gas}=0.7$.

Given the considerations of the relaxation parameters, one can obtain the desired bubble velocity
$U_{\mathrm{bubble}}$ from the equation for the Reynolds number (\ref{definitions:numbers}).
Substituting the bubble velocity $U_{\mathrm{bubble}}$ to the definition of the capillary number,
Eq. (\ref{definitions:numbers}), one can obtain the interface tension which is connected with the
binary liquid parameters $k$ and $A$. The stability of the lattice Boltzmann scheme implies the
simultaneous change of $\tau_{\phi}$ with the change of parameters $k$ and $A$
\cite{pagonabarraga-parameters}. Therefore, the recommendation is if Reynolds number is less than
$50-100$ one can neglect the inertia effects and obtain the bubble velocity straight from the
capillary number limiting the complex interplay of the parameters for the LBM simulations to be
stable: 
\begin{equation}
\begin{aligned}
&U_{\mathrm{bubble}}=Ca \frac{\gamma}{\mu_{\mathrm{liq}}}=Ca \frac{\sqrt{\frac{8 k
A}{9}}}{\frac{1}{3}(\tau_{\mathrm{liq}}-\frac{1}{2})}\\
\end{aligned}
\end{equation}
where usually $k=A=0.004,0.04$.

\item[Body force] 
It is desired to obtain the prescribed velocity $U_{\mathrm{bubble}}$ in simulations. The
flow in the simulations is conducted by imposing the body force. Thus, the assumption is needed of
how the body force is connected with the bubble velocity $U_{\mathrm{bubble}}$. We assume that the
flow is close to planar Poiseuille flow with $U_{\mathrm{bubble}}$ being the maximum in the
Poiseuille profile for the square shaped microchannels:
\begin{equation}
\label{poiseuille:square:profile}
\dfrac{\mathrm{d}P}{\mathrm{d}z}=\frac{\mu_{\mathrm{liq}} U_{\mathrm{bubble}}}{\sum_{i\leq 0,j\leq
0}{\dfrac{16}{\pi^6} \dfrac{(-1)^i (-1)^j}{(2 i+1)(2 j+1)\bigl[\frac{(2 i
+1)^2}{W_{\mathrm{eff}}}+\frac{(2 j+1)^2}{H_{\mathrm{eff}^2}}\bigr]}}}.
\end{equation}
Though the Poiseuille profile assumption works reasonably well \cite{kuzmin-binary2d}, instead of
calculating sums (\ref{poiseuille:square:profile})  one is suggested to start simulations with the
body force $\frac{\mathrm{d}P}{\mathrm{d}z}=10^{-6}-10^{-5}$ for grids of size $50\times 50 \times
1500$ and the capillary number $Ca\approx0.1-1.0$.  
\end{description}
After imposing the body force the simulations can be run. A typical simulation runs approximately
$100000-300000$ time steps to reach steady-state, see Section \ref{sec:steady:state}. To design the
simulations one needs to keep proportianality between parameters through equations above. For
example if one knows the capillary number $Ca_1$ and the corresponding body force
$\frac{\mathrm{d}P_1}{\mathrm{d}z}$ for simulations conducted on the grid with the characteristic
number of nodes $N_y^1$ , the body force for another simulation can be calculated as:
\begin{equation}
\dfrac{\dfrac{\mathrm{d}P_1}{\mathrm{d}x}}{\dfrac{\mathrm{d}P_2}{\mathrm{d}x}}=\frac{Ca_1}{Ca_2}
\frac{N_{
y2}^2 }{N_{y1}^2}
\end{equation}

\section{Symmetric boundary conditions}
\label{append:sym}
To reduce a computational load in terms of memory and CPU power we performed the simulation of the
quarter of the channel using the antisymmetric boundary conditions. In the macroscopic variables the
mirror boundary conditions require to have same scalar fields. For velocity boundary conditions only
tangential velocities are the same, perpendicular to boundary velocities have opposite
signs. This implies the following in terms of the
macroscopic variables:
\begin{equation}
\rho_B = \rho_F, \quad \phi_B = \phi_F \quad U_{B\tau}=U_{F\tau}, U_{B n}=U_{F n}, 
\end{equation}
where subscripts $B$ and $F$ stand for the boundary and the nearest fluid node, respectively.

For the lattice Boltzmann model the incoming populations on the boundary node need to be specified.
For generality we assume that the plane of the symmetry $\alpha=0$, where $\alpha$ is the
direction $x$ or $y$. Then, for lattice Boltzmann populations the symmetric boundary conditions can
be expressed through the following steps:
\begin{description}
\item[I] The first step is to find the complimentary directions. The complimentary $j$-th direction
to the $i$-th direction should satisfy three conditions:
\begin{equation}
\begin{aligned}
&c_{i\alpha}=-c_{j\alpha}, \text{ where $\alpha=$ is the symmetry plane }\\
&c_{i\beta}=-c_{j\beta}, \text{ where $\beta\neq\alpha$ }
\end{aligned}
\end{equation}

Note that the complimentary directions for the directions parallel to the symmetry plane coincide
with the original direction.

\item[II] The second step is taking the distribution populations from the nearest fluid node and
apply them to the boundary node utilizing complimentary directions:
\begin{equation}
f_{i,B}=f_{j,F},
\end{equation}
where $j$ is the index of the direction complimentary to the $i$-th direction.
\end{description}

By applying the procedure described above, one can obtain that the scalar fields and the tangential
velocity are conserved, the normal flux is reversed. In terms of the lattice Boltzmann populations
this condition improve the performance and memory requirements by a factor of $4$.

\bibliographystyle{unsrtnat}
\bibliography{paper}
\end{document}